### 2019-01-18 paper/paper.tex

%\def\useacmstyle{}
\def\submissionpapernumber{$\infty$}

\ifx\useacmstyle\undefined
    \documentclass[11pt]{article}
    \usepackage{fullpage}
    \usepackage[hidelinks]{hyperref}
\else
    \documentclass[sigconf,anonymous]{acmart}
    \fancyhf{}
    \fancyhead[C]{Anonymous submission \submissionpapernumber\ to ACM CCS 2019}
    \fancyfoot[C]{\thepage}
    \setcopyright{none}
    \acmConference{ACM CCS Submission}{2019 Feb 05}{}
    \acmYear{2019}
    \settopmatter{printacmref=false, printccs=false, printfolios=false}
\fi

\usepackage[dvipsnames]{xcolor}
\usepackage{amsthm,amsmath,amsfonts,amssymb,amstext,mathtools}
\usepackage{macros}
\usepackage[normalem]{ulem}
\usepackage{nicefrac}
\usepackage{peanutgallery}

\title{An Airdrop That Preserves Recipient Privacy}
\author{Dan Boneh \and Chris Jeffrey \and Joseph Poon \and Riad Wahby}

\begin{document}

\maketitle

\input{abstract}

\input{intro}

\input{defs}

\input{ecdsa.tex}

\input{rsa.tex}

\bibliographystyle{alpha}
\bibliography{paper}

\end{document}