### 2019-01-19 paper/intro.tex
move some stuff from intro to 'prac'

\section{Introduction}\label{sec:intro}

More to come.